---
title: "DS"
author: "Michael Chung"
date: "02/11/2021"
output: 
  html_document:
    df_print: kable
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
#Load packages/Set up environment
```{r message=FALSE, warning=FALSE}
#Load all packages used this semester just for fun
library(dplyr)
library(tidyverse)
library(performance)
library(LambertW)
library(patchwork)
library(rstan)
library(StanHeaders)
library(DHARMa)
library(fitdistrplus)
library(gamlss)
library(mgcv)
library(goft)
library(ggplot2)
library(brms)
library(modelsummary)
```

#Load data
```{r message=FALSE, warning=FALSE}
#Load dataset 
data = read.csv("Raw_data.csv")
```

#Data cleaning
```{r message=FALSE, warning=FALSE}
#Label treatment numbers as descriptions of levels of amputation 
data = data %>%mutate(Treatment = case_when(
    Treatment == "1" ~ "Control",
    Treatment == "2" ~ "Intermediate",
    Treatment == "3" ~ "Extreme"
  ))
```


#Now lets check the distribution of data and normalize it if necessary
```{r message=FALSE, warning=FALSE}

#Quick visualization of the response variable data
hist(data$Average_growth_cm)

#Figure out what distribution family the response variable fits best
fitDist(Average_growth_cm, data = data, type = "realAll", try.gamlss = T)

#Because it is in the Log distribution family,
#log transform the data to normalize it
logdata = data%>%
  mutate(Log_avg_growth = log10(Average_growth_cm))

#Check for normality, then for model assumptions using check model, if normal, use lm 
shapiro.test(logdata$Log_avg_growth)


#Data is still not normal -> Gaussianize
#To Gaussianize the data, first prepare it in a readable format
test <- data %>% 
  na.omit() %>%
  mutate(Average_growth_cm = as.numeric(Average_growth_cm), Individual = as.factor(Individual), Plot = as.factor(Plot))
#Now let us Gaussianize it 
test2 = Gaussianize(test$Average_growth_cm)
#Check to see whether or not Gaussianize worked to normalize the data 
shapiro.test(test2)
#Hooray it worked! Combine the Gaussianized data, rename the newly formed column
test3 <- cbind(test,test2) %>% rename(norm_avg_growth = 'Y1.X')
```

#Find the sample sizes for the methods and figure captions 
```{r}

#find sample sizes for damage ratings for methods and figures
test3 %>% count(Damage_rating_5_point_scale)
#find sample sizes for treatments for methods and figures
test3 %>% count(Treatment)
```

#Find average growth of groups
```{r}
#Get average growth by treatment 
average_growth_treatments = data%>%
  group_by(as.factor(Treatment)) %>%
  summarise(avg_GT = mean(Average_growth_cm, na.rm = T))
average_growth_treatments

#Get average growth by damage rating 
average_growth_damage = data%>%
  group_by(as.factor(Damage_rating_5_point_scale)) %>%
  summarise(avg_GD = mean(Average_growth_cm, na.rm = T))
average_growth_damage

```


#Great! Time to model this thing
```{r message=FALSE, warning=FALSE}
#Create a full gamlss model with all the inputs columns recorded using test3 as the data set
#Gaussianized data (norm_avg_growth) as the response variable
#Use plot and individuals as random variables

growth.full.gam <- gamlss(norm_avg_growth ~  Per_amputated_length + Per_amputated_mass + Length_remaining_cm  + Mass_remaining_g + Amputated_mass_g + Amputated_length_cm +  Initial_mass_g + Treatment + Initial_length_cm + Holdfast_presence_absence + Treatment:Damage_rating_5_point_scale + Damage_rating_5_point_scale + random(Individual) + random(Plot), data = test3, control = gamlss.control(n.cyc = 150))

summary(growth.full.gam)

#Use a backwards stepwise approach to reduce the model
backwards.full.gam <- stepGAIC(growth.full.gam, direction = "backward", trace = F)

#See the reduced formula and its results
formula(backwards.full.gam)
summary(backwards.full.gam)

#what does the reduced gamlss model look like as an anova?
z = aov(norm_avg_growth ~ Per_amputated_length + Per_amputated_mass + 
    Length_remaining_cm + Treatment + Damage_rating_5_point_scale + 
    random(Individual) + random(Plot), data = test3)
summary(z)
```

#What are the R-squared values of the significant continuous inputs in relation to growth?
```{r}
#use the cor function with norm_avg_growth as the response to find R-squared values
cor(test3$norm_avg_growth, test3$Per_amputated_mass)
cor(test3$norm_avg_growth, test3$Per_amputated_length)
cor(test3$norm_avg_growth, test3$Length_remaining_cm)
cor(test3$norm_avg_growth, test3$Damage_rating_5_point_scale)
```

#This only compared treatments to the control, what about an intermediate-extreme treatment comparison?
#What about interactive effects between damage and treatment?
```{r message=FALSE, warning=FALSE}
#To see if there are differences between extreme and intermediate treatments, run a TukeyHSD
x = aov(norm_avg_growth ~ as.factor(Treatment), data = test3)
summary(x)
TukeyHSD(x)
#not very helpful tbh because there are different assumptions from the gamlss model 
```

#Let's visualize this now and save the graphs, starting with a boxplot of the treatments 
```{r message=FALSE, warning=FALSE}
#Graph raw data to see trends 

#Treatment boxplot
Treatment_boxplot = ggplot(data = na.omit(data), aes(x  = as.factor(Treatment), y = Average_growth_cm, colour = Treatment, fill = Treatment))+
  geom_boxplot(alpha = 0.3)+
  scale_x_discrete(limits = c("Control", "Intermediate", "Extreme"))+
  labs(x = "Treatment", y  = "Average growth (cm)")+
  theme_classic()+ 
  theme(legend.position = "none")+
  geom_point(aes(colour = Treatment))

#Save it
ggsave(plot = Treatment_boxplot, filename = "Fig2.jpg")
```

#What about a damage rating boxplot
```{r message=FALSE, warning=FALSE}
#Damage rating boxplot

#change colours to create a gradient of health to represent the damage ratings
#Let green be healthy, and red being the most damaged
my_cols = c("0" = "green2", "1" = "gold2", "2" = "orange2", "3" = "salmon2", "4" = "red2") 

#Plot it
Damage_boxplot = ggplot(data = na.omit(data),
          aes(x = as.factor(Damage_rating_5_point_scale), 
              y = Average_growth_cm,
              colour = as.factor(Damage_rating_5_point_scale),
              fill = as.factor(Damage_rating_5_point_scale)
              ))+
  geom_point(aes(colour = as.factor(Damage_rating_5_point_scale)))+
  geom_boxplot(alpha = 0.3)+
  scale_x_discrete(limits = c("0", "1", "2","3","4"))+
  labs(x = "Damage rating", y  = "Average growth (cm)")+
  theme_classic()+ 
  theme(legend.position = "none")+
  scale_colour_manual(values = my_cols) +
  scale_fill_manual(values = my_cols)


#Save it
ggsave(plot = Damage_boxplot, filename = "Fig3.jpg")
```

#Length remaining had a significant effect on growth  in the reduced model, let's see what it looks like in a scatterplot
```{r message=FALSE, warning=FALSE}
#Length remaining vs. Average growth graph
Length_remaining_scatterplot = ggplot()+
  geom_point(data = na.omit(data), aes(x  = Length_remaining_cm, y = Average_growth_cm))+
  labs(x = "Length remaining (cm)", y = "Average growth (cm)")+
  geom_smooth(method = "lm", data = na.omit(data), aes(x = Length_remaining_cm, y = Average_growth_cm), colour = "red2")+
  theme_classic()
#Save it
ggsave(plot = Length_remaining_scatterplot, filename = "Fig4.jpg")
```

#Let's see what percent amputated mass and percent amputated length look like side-by-side
```{r message=FALSE, warning=FALSE}
#Percent amputation vs. Average growth graph

Per.l.amp.scatterplot = ggplot()+
  geom_point(data = na.omit(data), aes(x  = Per_amputated_length, y = Average_growth_cm))+
  labs(x = "Percent of amputated length", y = "Average growth (cm)")+
  geom_smooth(method = "lm", data = na.omit(data), aes(x = Per_amputated_length, y = Average_growth_cm))+
  theme_classic()

Per.m.amp.scatterplot = ggplot()+
  geom_point(data = na.omit(data), aes(x  = Per_amputated_mass, y = Average_growth_cm))+
  labs(x = "Percent of amputated mass", y = "Average growth (cm)")+
  geom_smooth(method = "lm", data = na.omit(data), aes(x = Per_amputated_mass, y = Average_growth_cm))+
  theme_classic()

Percent.amp.growth = Per.l.amp.scatterplot+Per.m.amp.scatterplot

ggsave(plot = Percent.amp.growth, filename = "Fig5.jpg")


#see collinearity of percent amputations
cor(data$Per_amputated_length, data$Per_amputated_mass)
```

#That's all folks! We got to the end!
